```{r}
#Load necessary libraries

library(tidyverse)
library(broom)
library(dplyr)

#Read in files
covid_data_cap <- read.csv("/Users/lorenzocapitani/OneDrive\ -\ Cardiff\ University/Bioinformatics/MartinPaper/April2022_Cap.csv")
covid_data_vein <- read.csv("/Users/lorenzocapitani/OneDrive\ -\ Cardiff\ University/Bioinformatics/MartinPaper/April2022_Vein.csv")

#Drop unneeded columns
covid_data_cap <- covid_data_cap[,-c(6,7,16:21)]
covid_data_vein <- covid_data_vein[,-c(4,6,8,10)]
```

```{r, fig.height=10}
#Clean dataset: convert binary factors to 1 or 0 
covid_data_cap$COVID.19.Positive. <- ifelse(test=covid_data_cap$COVID.19.Positive. == "Y", yes=1, no=0)
covid_data_cap$Vaccinated. <- ifelse(test=covid_data_cap$Vaccinated. == "Y", yes=1, no=0)
covid_data_cap$Prior.COVID.19. <- ifelse(test=covid_data_cap$Prior.COVID.19. == "Y", yes=1, no=0)
covid_data_cap[covid_data_cap$Gender == "male",]$Gender <- 0
covid_data_cap[covid_data_cap$Gender == "female",]$Gender <- 1
covid_data_cap$Gender <- as.integer(covid_data_cap$Gender)
covid_data_cap$Significant.co.morbidity. <- ifelse(test=covid_data_cap$Significant.co.morbidity. == "Y", yes=1, no=0)
covid_data_cap$Ethnicity..Y...BAME. <- ifelse(test=covid_data_cap$Ethnicity..Y...BAME. == "Y", yes=1, no=0)

#Repeat for venous dataset
covid_data_vein$COVID.19.Positive. <- ifelse(test=covid_data_vein$COVID.19.Positive. == "Y", yes=1, no=0)
covid_data_vein$Vaccinated. <- ifelse(test=covid_data_vein$Vaccinated. == "Y", yes=1, no=0)
covid_data_vein$Prior.COVID.19. <- ifelse(test=covid_data_vein$Prior.COVID.19. == "Y", yes=1, no=0)
covid_data_vein[covid_data_vein$Gender == "M",]$Gender <- 0
covid_data_vein[covid_data_vein$Gender == "F",]$Gender <- 1
covid_data_vein$Gender <- as.integer(covid_data_vein$Gender)
```


Correlation plotting before mean interpolation
Correlation matrices from Figures 1 and 2

```{r, fig.height=10, fig.width=10}
library(corrplot)

# Which parameters to include
numeric_covid_cap <- covid_data_cap[,c(2,3,4,5,6,7,8,9,10,11,12,13)]
numeric_covid_vein <- covid_data_vein[,c(2,3,4,5,6,7,8,9,10,11)]
covid_vein_cor_matrix <- cor(numeric_covid_vein, use = "complete.obs", method = "spearman")
covid_cap_cor_matrix <- cor(numeric_covid_cap, use = "complete.obs", method = "spearman")

#Obtain p-value matrix for each comparison and fix for multiple comparisons 
testRes_vein = cor.mtest(numeric_covid_vein, conf.level = 0.95,  method = "spearman", adjust="holm", exact = F)
testRes_cap = cor.mtest(numeric_covid_cap, conf.level = 0.95,  method = "spearman", adjust="holm", exact = F)

corrplot(covid_vein_cor_matrix, p.mat = testRes_vein$p, sig.level = 0.05, method = 'square', order="original", tl.cex = 2, tl.col = "black", insig = "blank", type = 'lower') 
title(main = "Vein dataset", cex.main =2, line = 0)
corrplot(covid_cap_cor_matrix, p.mat = testRes_cap$p, sig.level = 0.05, method = 'square', order="original", tl.cex = 2, tl.col = "black", insig = "blank", type = 'lower') 
title(main = "Capillary dataset", cex.main =2, line = 0)

```

Mean interpolation of few missing data points cprior to modelling

```{r, fig.height=10}
#A visual take on the missing values 
library(Amelia)
missmap(covid_data_vein, main = "Missing values vs observed in vein dataset")
missmap(covid_data_cap, main = "Missing values vs observed in cap dataset")

# Mean interpolation for the few missing values present for vein dataset
covid_data_vein$RBD.IgG..BAU.ml.[is.na(covid_data_vein$RBD.IgG..BAU.ml.)] <- mean(covid_data_vein$RBD.IgG..BAU.ml.,na.rm=T)
covid_data_vein$S1.IgG..BAU.ml.[is.na(covid_data_vein$S1.IgG..BAU.ml.)] <- mean(covid_data_vein$S1.IgG..BAU.ml., na.rm=T)
covid_data_vein$S2.IgG..BAU.ml.[is.na(covid_data_vein$S2.IgG..BAU.ml.)] <- mean(covid_data_vein$S2.IgG..BAU.ml., na.rm = T)
covid_data_vein$N.IgG..BAU.ml.[is.na(covid_data_vein$N.IgG..BAU.ml.)] <- mean(covid_data_vein$N.IgG..BAU.ml., na.rm=T)

# Mean interpolation for the few missing values present for cap dataset
covid_data_cap$Age[is.na(covid_data_cap$Age)] <- mean(covid_data_cap$Age,na.rm=T)
covid_data_cap$IFNG[is.na(covid_data_cap$IFNG)] <- mean(covid_data_cap$IFNG,na.rm=T)
covid_data_cap$RBD.IgG..BAU.ml.[is.na(covid_data_cap$RBD.IgG..BAU.ml.)] <- mean(covid_data_cap$RBD.IgG..BAU.ml., na.rm = T)
covid_data_cap$N.IgG..BAU.ml.[is.na(covid_data_cap$N.IgG..BAU.ml.)] <- mean(covid_data_cap$N.IgG..BAU.ml., na.rm = T)
covid_data_cap$S1.IgG..BAU.ml.[is.na(covid_data_cap$S1.IgG..BAU.ml.)] <- mean(covid_data_cap$S1.IgG..BAU.ml., na.rm = T)
covid_data_cap$S2.IgG..BAU.ml.[is.na(covid_data_cap$S2.IgG..BAU.ml.)] <- mean(covid_data_cap$S2.IgG..BAU.ml., na.rm=T)
# Drop row with missing gender value
covid_data_cap <- covid_data_cap[!is.na(covid_data_cap$Gender),]

missmap(covid_data_vein, main = "Missing values vs observed in vein dataset")
missmap(covid_data_cap, main = "Missing values vs observed in cap dataset")

```


Preliminary univariate analysis of venous dataset - not included in study

```{r}
library(ggplot2)
library(ggbeeswarm)
library(ggpubr)

numeric_covid_vein <- covid_data_vein[,c(2,3,4,5,6,7,8,9,10,11)]
numeric_covid_vein$COVID.19.Positive. <- factor(numeric_covid_vein$COVID.19.Positive.)
numeric_covid_vein$Vaccinated. <- factor(numeric_covid_vein$Vaccinated.)
numeric_covid_vein$Prior.COVID.19. <- factor(numeric_covid_vein$Prior.COVID.19.)
numeric_covid_vein$Gender <- factor(numeric_covid_vein$Gender)


ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., y = IFNg)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("IFNg (pg/ml)") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., y = RBD.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-RBD IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., y = N.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-RBD IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., y = S1.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-S1 IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., y = S2.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-S2 IgG titres") +
  stat_compare_means(label.x.npc = 0.4)


ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., fill = Vaccinated.)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Vaccinated") +
  scale_fill_manual(labels = c("No", "Yes"), values = c("brown3", "darkgoldenrod1")) 


ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., fill = Prior.COVID.19.)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Prior COVID infection") +
  scale_fill_manual(labels = c("No", "Yes"), values = c("brown3", "darkgoldenrod1")) 

ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., fill = Gender)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Gender") +
  scale_fill_manual(labels = c("Male", "Female"), values = c("brown3", "darkgoldenrod1")) 

ggplot(numeric_covid_vein, aes(x = COVID.19.Positive., y = Age)) +
    geom_violin() +
  geom_quasirandom(width = 0.2) +
  xlab("Breakthrough COVID19 infection") +
  ylab("Age") +
    scale_x_discrete(labels = c("No", "Yes"))+
  stat_compare_means(label.x.npc = 0.4)

```


Preliminary univariate analysis of capillary dataset - not included in study

```{r}

numeric_covid_cap <-  covid_data_cap[,c(2,3,4,5,6,7,8,9,10,11,12,13)]
numeric_covid_cap$COVID.19.Positive. <- factor(numeric_covid_cap $COVID.19.Positive.)
numeric_covid_cap$Vaccinated. <- factor(numeric_covid_cap $Vaccinated.)
numeric_covid_cap$Prior.COVID.19. <- factor(numeric_covid_cap $Prior.COVID.19.)
numeric_covid_cap$Gender <- factor(numeric_covid_cap $Gender)
numeric_covid_cap$Significant.co.morbidity. <- factor(numeric_covid_cap$Significant.co.morbidity.)
numeric_covid_cap$Ethnicity..Y...BAME. <- factor(numeric_covid_cap$Ethnicity..Y...BAME.)
ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., y = IFNG)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("IFNg (pg/ml)") +stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_cap, aes(x = COVID.19.Positive., y = RBD.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-RBD IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_cap, aes(x = COVID.19.Positive., y = N.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-RBD IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_cap, aes(x = COVID.19.Positive., y = S1.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-S1 IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_cap, aes(x = COVID.19.Positive., y = S2.IgG..BAU.ml.)) +
  geom_violin() +
  geom_quasirandom(width = 0.2) +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Ant-S2 IgG titres") +
  stat_compare_means(label.x.npc = 0.4)

ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., fill = Vaccinated.)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Vaccinated") +
  scale_fill_manual(labels = c("No", "Yes"), values = c("brown3", "darkgoldenrod1")) 



ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., fill = Prior.COVID.19.)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Prior COVID infection") +
  scale_fill_manual(labels = c("No", "Yes"), values = c("brown3", "darkgoldenrod1")) 

ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., fill = Gender)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Gender") +
  scale_fill_manual(labels = c("Male", "Female"), values = c("brown3", "darkgoldenrod1")) 

ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., fill = Significant.co.morbidity.)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Significant comorbidity") +
  scale_fill_manual(labels = c("No", "Yes"), values = c("brown3", "darkgoldenrod1")) 


ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., y = Age)) +
    geom_violin() +
  geom_quasirandom(width = 0.2) +
  xlab("Breakthrough COVID19 infection") +
  ylab("Age") +
    scale_x_discrete(labels = c("No", "Yes"))+
  stat_compare_means(label.x.npc = 0.4)


ggplot(numeric_covid_cap , aes(x = COVID.19.Positive., fill = Ethnicity..Y...BAME.)) +
  geom_bar(stat ='count',position = "dodge2") +
  scale_x_discrete(labels = c("No", "Yes")) + 
  xlab("Breakthrough COVID19 infection") +
  ylab("Count") +
  labs(fill = "Ethnicity") +
  scale_fill_manual(labels = c("Non-BAME", "BAME"), values = c("brown3", "darkgoldenrod1")) 

```


Logistic regression modelling for venous dataset used in Extended Figure 3

```{r}
library(caret)
library(OddsPlotty)
library(ggplot2)
library(OddsPlotty)
library(caret)
library(tidyverse)
library(rms)

numeric_covid_vein <- covid_data_vein[,c(2,3,4,5,6,7,8,9,10,11)]
#Conversion of numeric measurements to quartiles

q1 <- summary(numeric_covid_vein$IFNg)[2]
q2 <- summary(numeric_covid_vein$IFNg)[3]
q3 <- summary(numeric_covid_vein$IFNg)[5]

q1r <- summary(numeric_covid_vein$RBD.IgG..BAU.ml.)[2]
q2r <- summary(numeric_covid_vein$RBD.IgG..BAU.ml.)[3]
q3r <- summary(numeric_covid_vein$RBD.IgG..BAU.ml.)[5]

q1s1 <- summary(numeric_covid_vein$S1.IgG..BAU.ml.)[2]
q2s1<- summary(numeric_covid_vein$S1.IgG..BAU.ml.)[3]
q3s1 <- summary(numeric_covid_vein$S1.IgG..BAU.ml.)[5]

q1s2<- summary(numeric_covid_vein$S2.IgG..BAU.ml.)[2]
q2s2<- summary(numeric_covid_vein$S2.IgG..BAU.ml.)[3]
q3s2 <- summary(numeric_covid_vein$S2.IgG..BAU.ml.)[5]

q1n <- summary(numeric_covid_vein$N.IgG..BAU.ml.)[2]
q2n <- summary(numeric_covid_vein$N.IgG..BAU.ml.)[3]
q3n  <- summary(numeric_covid_vein$N.IgG..BAU.ml.)[5]

numeric_covid_vein[numeric_covid_vein$IFNg < q1,]$IFNg <- 1
numeric_covid_vein[numeric_covid_vein$IFNg >= q1 & numeric_covid_vein$IFNg < q2,]$IFNg <- 2
numeric_covid_vein[numeric_covid_vein$IFNg >= q2 & numeric_covid_vein$IFNg < q3,]$IFNg <- 3
numeric_covid_vein[numeric_covid_vein$IFNg >= q3,]$IFNg <- 4

numeric_covid_vein[numeric_covid_vein$RBD.IgG..BAU.ml. < q1r,]$RBD.IgG..BAU.ml. <- 1
numeric_covid_vein[numeric_covid_vein$RBD.IgG..BAU.ml. >= q1r & numeric_covid_vein$RBD.IgG..BAU.ml. < q2r,]$RBD.IgG..BAU.ml. <- 2
numeric_covid_vein[numeric_covid_vein$RBD.IgG..BAU.ml. >= q2r & numeric_covid_vein$RBD.IgG..BAU.ml. < q3r,]$RBD.IgG..BAU.ml. <- 3
numeric_covid_vein[numeric_covid_vein$RBD.IgG..BAU.ml. >= q3r,]$RBD.IgG..BAU.ml. <- 4

numeric_covid_vein[numeric_covid_vein$S1.IgG..BAU.ml. < q1s1,]$S1.IgG..BAU.ml. <- 1
numeric_covid_vein[numeric_covid_vein$S1.IgG..BAU.ml. >= q1s1 & numeric_covid_vein$S1.IgG..BAU.ml. < q2s1,]$S1.IgG..BAU.ml. <- 2
numeric_covid_vein[numeric_covid_vein$S1.IgG..BAU.ml. >= q2s1 & numeric_covid_vein$S1.IgG..BAU.ml. < q3s1,]$S1.IgG..BAU.ml. <- 3
numeric_covid_vein[numeric_covid_vein$S1.IgG..BAU.ml. >= q3s1,]$S1.IgG..BAU.ml. <- 4

numeric_covid_vein[numeric_covid_vein$S2.IgG..BAU.ml. < q1s2,]$S2.IgG..BAU.ml. <- 1
numeric_covid_vein[numeric_covid_vein$S2.IgG..BAU.ml. >= q1s2 & numeric_covid_vein$S2.IgG..BAU.ml. < q2s2,]$S2.IgG..BAU.ml. <- 2
numeric_covid_vein[numeric_covid_vein$S2.IgG..BAU.ml. >= q2s2 & numeric_covid_vein$S2.IgG..BAU.ml. < q3s2,]$S2.IgG..BAU.ml. <- 3
numeric_covid_vein[numeric_covid_vein$S2.IgG..BAU.ml. >= q3s2,]$S2.IgG..BAU.ml. <- 4

numeric_covid_vein[numeric_covid_vein$N.IgG..BAU.ml. < q1n,]$N.IgG..BAU.ml. <- 1
numeric_covid_vein[numeric_covid_vein$N.IgG..BAU.ml. >= q1n & numeric_covid_vein$N.IgG..BAU.ml. < q2n,]$N.IgG..BAU.ml. <- 2
numeric_covid_vein[numeric_covid_vein$N.IgG..BAU.ml. >= q2n & numeric_covid_vein$N.IgG..BAU.ml. < q3n,]$N.IgG..BAU.ml. <- 3
numeric_covid_vein[numeric_covid_vein$N.IgG..BAU.ml. >= q3n,]$N.IgG..BAU.ml. <- 4

numeric_covid_vein$IFNg <- factor(numeric_covid_vein$IFNg, levels = c(1,2,3,4))
numeric_covid_vein$RBD.IgG..BAU.ml. <- factor(numeric_covid_vein$RBD.IgG..BAU.ml., levels = c(1,2,3,4))
numeric_covid_vein$S1.IgG..BAU.ml. <- factor(numeric_covid_vein$S1.IgG..BAU.ml., levels = c(1,2,3,4))
numeric_covid_vein$S2.IgG..BAU.ml. <- factor(numeric_covid_vein$S2.IgG..BAU.ml., levels = c(1,2,3,4))
numeric_covid_vein$N.IgG..BAU.ml. <- factor(numeric_covid_vein$N.IgG..BAU.ml., levels = c(1,2,3,4))

#Logistic modelling

model_vein <- glm(COVID.19.Positive.~ ., data=numeric_covid_vein, family=binomial(link='logit'))

summary(model_vein)

#Extraction of odds ratios 
odds <- odds_plot(model_vein)
odds <- odds$odds_data
write.csv(odds, file = "...")
odds$OR <- log(odds$OR)
odds$lower <- log(odds$lower)
odds$upper <- log(odds$upper)

#Odds ratio plotting
ggplot(odds, aes(x = OR, y = vars)) + 
    geom_vline(aes(xintercept = 0), size = .25, linetype = "dashed") + 
    geom_errorbarh(aes(xmax = upper, xmin = lower), size = .5, height = 
                    .2, color = "gray50") +
    geom_point(size = 3.5, color = "orange") +
    theme_bw()+
    theme(panel.grid.minor = element_blank()) +
    ylab("") +
    xlab("log(Odds ratio)") +
    ggtitle("Parameters and the risk of COVID infections - Venous cohort") +
  scale_x_continuous(limits = c(-6,6))

# Odd as percentages
(exp(model_vein$coefficients[-1])-1)*100


```

Logistic regression modelling for capillary dataset used in Extended Figure 3


```{r}
numeric_covid_cap <- covid_data_cap[,c(2,3,4,5,6,7,8,9,10,11,12,13)]

q1 <- summary(numeric_covid_cap$IFNG)[2]
q2 <- summary(numeric_covid_cap$IFNG)[3]
q3 <- summary(numeric_covid_cap$IFNG)[5]

q1r <- summary(numeric_covid_cap$RBD.IgG..BAU.ml.)[2]
q2r <- summary(numeric_covid_cap$RBD.IgG..BAU.ml.)[3]
q3r <- summary(numeric_covid_cap$RBD.IgG..BAU.ml.)[5]

q1s1 <- summary(numeric_covid_cap$S1.IgG..BAU.ml.)[2]
q2s1<- summary(numeric_covid_cap$S1.IgG..BAU.ml.)[3]
q3s1 <- summary(numeric_covid_cap$S1.IgG..BAU.ml.)[5]

q1s2<- summary(numeric_covid_cap$S2.IgG..BAU.ml.)[2]
q2s2<- summary(numeric_covid_cap$S2.IgG..BAU.ml.)[3]
q3s2 <- summary(numeric_covid_cap$S2.IgG..BAU.ml.)[5]

q1n <- summary(numeric_covid_cap$N.IgG..BAU.ml.)[2]
q2n <- summary(numeric_covid_cap$N.IgG..BAU.ml.)[3]
q3n  <- summary(numeric_covid_cap$N.IgG..BAU.ml.)[5]

numeric_covid_cap[numeric_covid_cap$IFNG < q1,]$IFNG <- 1
numeric_covid_cap[numeric_covid_cap$IFNG >= q1 & numeric_covid_cap$IFNG < q2,]$IFNG <- 2
numeric_covid_cap[numeric_covid_cap$IFNG >= q2 & numeric_covid_cap$IFNG < q3,]$IFNG <- 3
numeric_covid_cap[numeric_covid_cap$IFNG >= q3,]$IFNG <- 4

numeric_covid_cap[numeric_covid_cap$RBD.IgG..BAU.ml. < q1r,]$RBD.IgG..BAU.ml. <- 1
numeric_covid_cap[numeric_covid_cap$RBD.IgG..BAU.ml. >= q1r & numeric_covid_cap$RBD.IgG..BAU.ml. < q2r,]$RBD.IgG..BAU.ml. <- 2
numeric_covid_cap[numeric_covid_cap$RBD.IgG..BAU.ml. >= q2r & numeric_covid_cap$RBD.IgG..BAU.ml. < q3r,]$RBD.IgG..BAU.ml. <- 3
numeric_covid_cap[numeric_covid_cap$RBD.IgG..BAU.ml. >= q3r,]$RBD.IgG..BAU.ml. <- 4

numeric_covid_cap[numeric_covid_cap$S1.IgG..BAU.ml. < q1s1,]$S1.IgG..BAU.ml. <- 1
numeric_covid_cap[numeric_covid_cap$S1.IgG..BAU.ml. >= q1s1 & numeric_covid_cap$S1.IgG..BAU.ml. < q2s1,]$S1.IgG..BAU.ml. <- 2
numeric_covid_cap[numeric_covid_cap$S1.IgG..BAU.ml. >= q2s1 & numeric_covid_cap$S1.IgG..BAU.ml. < q3s1,]$S1.IgG..BAU.ml. <- 3
numeric_covid_cap[numeric_covid_cap$S1.IgG..BAU.ml. >= q3s1,]$S1.IgG..BAU.ml. <- 4

numeric_covid_cap[numeric_covid_cap$S2.IgG..BAU.ml. < q1s2,]$S2.IgG..BAU.ml. <- 1
numeric_covid_cap[numeric_covid_cap$S2.IgG..BAU.ml. >= q1s2 & numeric_covid_cap$S2.IgG..BAU.ml. < q2s2,]$S2.IgG..BAU.ml. <- 2
numeric_covid_cap[numeric_covid_cap$S2.IgG..BAU.ml. >= q2s2 & numeric_covid_cap$S2.IgG..BAU.ml. < q3s2,]$S2.IgG..BAU.ml. <- 3
numeric_covid_cap[numeric_covid_cap$S2.IgG..BAU.ml. >= q3s2,]$S2.IgG..BAU.ml. <- 4

numeric_covid_cap[numeric_covid_cap$N.IgG..BAU.ml. < q1n,]$N.IgG..BAU.ml. <- 1
numeric_covid_cap[numeric_covid_cap$N.IgG..BAU.ml. >= q1n & numeric_covid_cap$N.IgG..BAU.ml. < q2n,]$N.IgG..BAU.ml. <- 2
numeric_covid_cap[numeric_covid_cap$N.IgG..BAU.ml. >= q2n & numeric_covid_cap$N.IgG..BAU.ml. < q3n,]$N.IgG..BAU.ml. <- 3
numeric_covid_cap[numeric_covid_cap$N.IgG..BAU.ml. >= q3n,]$N.IgG..BAU.ml. <- 4

numeric_covid_cap$IFNG <- factor(numeric_covid_cap$IFNG, levels = c(1,2,3,4))
numeric_covid_cap$RBD.IgG..BAU.ml. <- factor(numeric_covid_cap$RBD.IgG..BAU.ml., levels = c(1,2,3,4))
numeric_covid_cap$S1.IgG..BAU.ml. <- factor(numeric_covid_cap$S1.IgG..BAU.ml., levels = c(1,2,3,4))
numeric_covid_cap$S2.IgG..BAU.ml. <- factor(numeric_covid_cap$S2.IgG..BAU.ml., levels = c(1,2,3,4))
numeric_covid_cap$N.IgG..BAU.ml. <- factor(numeric_covid_cap$N.IgG..BAU.ml., levels = c(1,2,3,4))


model_cap <- glm(COVID.19.Positive.~ ., data=numeric_covid_cap, family=binomial(link='logit'))
summary(model_cap)

odds <- odds_plot(model_cap)
odds <- odds$odds_data
odds$OR <- log(odds$OR)
odds$lower <- log(odds$lower)
odds$upper <- log(odds$upper)


ggplot(odds, aes(x = OR, y = vars)) + 
    geom_vline(aes(xintercept = 0), size = .25, linetype = "dashed") + 
    geom_errorbarh(aes(xmax = upper, xmin = lower), size = .5, height = 
                    .2, color = "gray50") +
    geom_point(size = 3.5, color = "orange") +
    theme_bw()+
    theme(panel.grid.minor = element_blank()) +
    ylab("") +
    xlab("log(Odds ratio)") +
    ggtitle("Parameters and the risk of COVID infections - capillary dataset") +
  scale_x_continuous(limits = c(-10,1))

# Odd as percentages
(exp(model_cap$coefficients[-1])-1)*100

```

Cross-validation of logisitc regression model performed on venous dataset using bestglm

```{r}
library(bestglm)
#numeric_covid_vein 

numeric_covid_vein$COVID.19.Positive. <- factor(numeric_covid_vein$COVID.19.Positive., levels = c(0,1))


test_bestglm <-numeric_covid_vein[, c(names(numeric_covid_vein)[-1], "COVID.19.Positive.")]
names(test_bestglm)[ncol(test_bestglm)] <- "y"

test_bestglm$Gender <- as.numeric(test_bestglm$Gender)
test_bestglm$IFNg <- as.factor(test_bestglm$IFNg)
test_bestglm$RBD.IgG..BAU.ml. <- as.factor(test_bestglm$RBD.IgG..BAU.ml.)
test_bestglm$S1.IgG..BAU.ml. <- as.factor(test_bestglm$S1.IgG..BAU.ml.)
test_bestglm$S2.IgG..BAU.ml. <- as.factor(test_bestglm$S2.IgG..BAU.ml.)
test_bestglm$N.IgG..BAU.ml. <- as.factor(test_bestglm$N.IgG..BAU.ml.)

test_bestglm <-
    bestglm(Xy = test_bestglm,
            family = binomial,
            IC = "AIC",                 # Information criteria for
            method = "exhaustive")


summary(test_bestglm$BestModel)

odds <- odds_plot(test_bestglm$BestModel)
odds <- odds$odds_data
write.csv(odds, file = "/Users/lorenzocapitani/OneDrive\ -\ Cardiff\ University/Bioinformatics/MartinPaper/odds_final.csv")
odds$OR <- log(odds$OR)
odds$lower <- log(odds$lower)
odds$upper <- log(odds$upper)


ggplot(odds, aes(x = OR, y = vars)) + 
    geom_vline(aes(xintercept = 0), size = .25, linetype = "dashed") + 
    geom_errorbarh(aes(xmax = upper, xmin = lower), size = .5, height = 
                    .2, color = "gray50") +
    geom_point(size = 3.5, color = "orange") +
    theme_bw()+
    theme(panel.grid.minor = element_blank()) +
    ylab("") +
    xlab("log(Odds ratio)") +
    ggtitle("Parameters and the risk of COVID infections - Venous cohort") +
  scale_x_continuous(limits = c(-6,6))

# Odd as percentages
(exp(test_bestglm$BestModel$coefficients[-1])-1)*100

```
